Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-30 manuscript/HIAs.qmd
validate

@@

>>>>

```{r rescale}
#| eval: true

# Now I rescale the now continous variables using reference and threshold values
# I will use the same reference levels/values for all of Norway:

upper <- 0
lower <- 100
threshold <- 10

eaTools::ea_normalise(data = naturetypes,
                      vector = "ADSV",
                      upper_reference_level = lower,
                      lower_reference_level = upper,
                      break_point = threshold,
                      plot=T,
                      reverse = T
                      )

# There is no point yet making this a time series
# I will assign all the indicator value to the same time (2018-2022)

# same for 7FK
eaTools::ea_normalise(data = naturetypes,
                      vector = "var_7FA",
                      upper_reference_level = lower,
                      lower_reference_level = upper,
                      break_point = threshold,
                      plot=T,
                      reverse = T
                      )
# The variables are really coarse
```

# Adding scaled indicator values to the dataset
naturetypes$i_ADSV <- eaTools::ea_normalise(data = naturetypes,
                      vector = "ADSV",
                      upper_reference_level = lower,
                      lower_reference_level = upper,
                      break_point = threshold,
                      reverse = T
                      )

naturetypes$i_alien <- eaTools::ea_normalise(data = naturetypes,
                      vector = "var_7FA",
                      upper_reference_level = lower,
                      lower_reference_level = upper,
                      break_point = threshold,
                      reverse = T
                      )
```

@@

```{r HIA-validate}
#| eval: false

# Now I want to see if the indicator values depend on the HIA is a predictable
# way to justify the stratification
infra |>
  as_tibble() |>
  count(infrastructureIndex)

corrCheck <- st_intersection(naturetypes, infra)
corrCheck |>
  as_tibble() |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  count(i_ADSV_fct, infrastructureIndex) |>
  View()

corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    #position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(~indicator)
```

```{r}
#| eval: false

# There is very little data from HIA 3, so I combine it with HIA 2
corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10,
         infrastructureIndex = case_when(
           infrastructureIndex == 3 ~ 2,
           .default = infrastructureIndex
         )) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(~indicator)
```

```{r}
# That did not made it better
# Is there a year effect?

corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10,
         infrastructureIndex = case_when(
           infrastructureIndex == 3 ~ 2,
           .default = infrastructureIndex
         )) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(indicator~year)
```

```{r}
# Next, make a violing/box plot

corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10,
         infrastructureIndex = case_when(
           infrastructureIndex == 3 ~ 2,
           .default = infrastructureIndex
         )) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    y = indicatorValue))+
  geom_boxplot(
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(indicator~year)
```

```{r}
#| eval: false

# All the 1's might make it hard to see the pattern.
# Try removing the 1's (the areas in best condition)

corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10,
         infrastructureIndex = case_when(
           infrastructureIndex == 3 ~ 2,
           .default = infrastructureIndex
           )) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  mutate(indicatorValue = na_if(indicatorValue, 1)) |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(~indicator)
